---
title: "Reports on the US gun murders"
author: "Fereidoun Mianji"
date: "10/25/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This is a report of 2010 gun murder rates obtained from FBI reports. The data was taken from this url:
"https://raw.githubusercontent.com/rafalab/dslabs/master/inst/
extdata/murders.csv" that is wrangled by Rafael Irizarry from the original data available on (https://en.wikipedia.org/wiki/Murder_in_the_united_states_by_state). The following library is used to analyze the data.

```{r loading libs, message=FALSE}
library(tidyverse)
```

## Analyzing the murder rate by state

Then the data is wrangled and saved as murders.rda.

```{r calculating the rates}
murders <- read_csv("data/murders.csv")
murders <-murders %>% mutate(region = factor(region),
                             rate = total / population * 10^5)
save(murders, file = "rdas/murders.rda")
```

## The plot

The barplot depicting the murder rates in all states was generated using the following code. It shows the significat variability of murder rates by state.

```{r plotting and saving the figure}
load("rdas/murders.rda")
murders %>% mutate(abb = reorder(abb, rate)) %>%
  ggplot(aes(abb, rate)) +
  geom_bar(width = 0.5, stat = "identity", color = "black") +
  coord_flip() +
  ggsave("figs/barplot.png")

```


